The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-03-01 analysis/paper/paper.qmd
Update geocode function and function call in paper

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                        # Managing paths
  "tidyverse",                                   # Data manipulation and visualisation
  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
  "tidygeocoder",                                # Working with spatial data
  "igraph", "ggraph"                             # Network analysis and visualsation
)

## Load required packages, install them if needed
for (package in packages) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}

@@

The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined. With the exception of a bilingual English-Spanish thesis, all thesis were written in English.

A total number of `r nrow(emu_theses)` out of `r nrow(theses_all)` were analysed.

@@

# Geocode thesis locations
emu_theses <- emu_theses %>%
  geocode(location, method = 'osm', lat = latitude , long = longitude)

# Show thesis locations on world map
world <- map_data("world")
ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray", size = 0.1
  ) +
  geom_count(data = emu_theses, aes(longitude, latitude), color = "red", alpha = 0.7) +
  scale_size_area(max_size = 3) +
  coord_fixed() +
  theme_void()
```

@@

```{r}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_theses_words, 
                           top_n = 3, 
                           title_col = "title", 
                           word_col = "word")
```

```{r}
# Which are the top n most frequently used words in all theses`?
get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 5, 
                         word_col = "word")
```

```{r}
# Which theses contain "fabric" in the top 20 most frequently used words?
emu_theses_top_20 %>% 
  filter(word == "fabric")

emu_theses_top_20 %>%
  slice(1:40) %>% 
  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2)
```

```{r}
#| label: ngrams

## TO DO: generalise bigrams and trigrams into ngrams function

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
  anti_join(stop_words, by = c("first" = "word")) |> 
  anti_join(stop_words, by = c("second" = "word")) |> 
  filter(str_detect(first, "[a-z]") & 
         str_detect(second, "[a-z]"))

bigram_counts <- emu_theses_bigrams |> 
  count(first, second, sort = TRUE)

bigram_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
  filter(!first %in% stop_words$word,
         !second %in% stop_words$word,
         !third %in% stop_words$word) |>  
  filter(str_detect(first, "[a-z]"), 
         str_detect(second, "[a-z]"),
         str_detect(third, "[a-z]"))

trigram_counts <- emu_theses_trigrams |> 
  count(first, second, sort = TRUE)

trigram_counts

## Bigram tf-idf
bigram_tf_idf <- emu_theses_bigrams |> 
  count(title, bigram) |> 
  bind_tf_idf(bigram, title, n) |> 
  arrange(desc(tf_idf))

bigram_tf_idf

## Trigram tf-idf
trigram_tf_idf <- emu_theses_trigrams |> 
  count(title, trigram) |> 
  bind_tf_idf(trigram, title, n) |> 
  arrange(desc(tf_idf))

trigram_tf_idf

## Show network of bigrams
bigram_graph <- bigram_counts |> 
  filter(n > 100) |> 
  graph_from_data_frame()

bigram_graph

set.seed(2023)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

## Higgh frequency bigrams ----

emu_theses_bigrams_top <- emu_theses_bigrams %>%
  group_by(title) %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() %>% 
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top %>%
  filter(title == unique(title)[1:6]) |> 
  ggplot(aes(x = bigram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

```{r}
#| label: tf-idf

emu_theses_words

emu_theses_words_count <- emu_theses_words |> 
  count(title, word, sort = TRUE) |> 
  ungroup()

emu_theses_words_total <- emu_theses_words_count |> 
  group_by(title) |> 
  summarise(total = sum(n))

emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)

# emu_theses_words_count |> 
#   group_by(title) |> 
#   summarise(total = unique(tot

# Preview term frequency distributions in a subset of theses
ggplot(emu_theses_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")

# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_theses_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

## Visualise Zipf's law
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  scale_x_log10() +
  scale_y_log10()

rank_subset <- freq_by_rank |> 
  filter(rank < 500,
         rank > 10)

lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()

## Calculate tf-idf
emu_theses_words_count <- emu_theses_words_count |> 
  bind_tf_idf(word, title, n)
emu_theses_words_count

emu_theses_words_count |> 
  select(-total) |> 
  arrange(desc(tf_idf))

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_theses_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words. As such, it shows words that are the most important to one document in a collection of documents [@silge2017].

>>>>

# Load packages ----
## Managing paths
library(here)

## Data manipulation and visualisation
library(tidyverse)

## Text processing and analysis
library(tidytext)
library(SnowballC)
library(tm)

## Working with spatial data
library(tidygeocoder)

@@

The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined.

A total number of `r nrow(emu_theses)` were analysed.

@@

# Add latitude and longitude information to theses data frame
emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)

# Show thesis locations on world map
visualize_thesis_locations(emu_theses)
```

@@

```{r}
# Which are the top 20 most frequently used words in each thesis?
emu_theses_top_20 <- emu_theses_words %>%
  group_by(title) %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50) %>%              ### filter words with a minimum count
  ungroup() %>% 
  mutate(word = reorder(word, n))

# Which are first 5 most frequently used words in each thesis?
emu_theses_first_5 <- emu_theses_words %>%
  count(title, word) %>% 
  group_by(title) %>% 
  arrange(desc(n)) %>% 
  slice(1:5)

# Which theses contain "fabric" in the top 20 most frequently used words?
emu_theses_top_20 %>% 
  filter(word == "fabric")

emu_theses_top_20 %>%
  slice(1:40) %>% 
  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2)

# # wip function ngrams
# get_ngrams <- function(df, n) {
#   if (!is.numeric(n)) {
#     stop("n is not numeric")
#   }
#   emu_theses %>%
#     mutate(author = paste(first_name, last_name)) %>% 
#     select(author, title, text, -c(year, first_name, last_name, text_start, text_end)) %>% 
#     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
#     separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
#     anti_join(stop_words, by = c("first" = "word")) %>%
#     anti_join(stop_words, by = c("second" = "word")) %>%
#     anti_join(stop_words_custom, by = c("first" = "word")) %>%
#     anti_join(stop_words_custom, by = c("second" = "word")) %>%
#     filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
#     filter(first != second) %>% 
#     filter(first != "source")
# }
# emu_theses_ngrams <- get_ngrams(emu_theses, 2)

# bigrams
emu_theses_bigrams <- emu_theses %>%
  mutate(author = paste(first_name, last_name)) %>% 
  select(author, title, text, -c(graduation_year, first_name, last_name)) %>% 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
  separate(bigram, into = c("first","second"), sep = " ", remove = FALSE) %>%
  anti_join(stop_words, by = c("first" = "word")) %>%
  anti_join(stop_words, by = c("second" = "word")) %>%
  anti_join(stop_words_custom, by = c("first" = "word")) %>%
  anti_join(stop_words_custom, by = c("second" = "word")) %>%
  filter(str_detect(first, "[a-z]") & str_detect(second, "[a-z]")) %>% 
  filter(first != second) %>% 
  filter(first != "source")

emu_theses_bigrams_top <- emu_theses_bigrams %>%
  group_by(title) %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 10) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() %>% 
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top %>%
  ggplot(aes(x = bigram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2)

# find the most frequently used bigrams
emu_thesis_bigrams_top <- emu_theses_bigrams %>%
  group_by(title) %>% 
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 10) %>%                   ### change the number of top words to show
  mutate(bigram = reorder(bigram, n)) %>% 
  print()
```
